Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-10 paper/paper.tex
Updated

\subsection{A liveness-based garbage collection scheme}

  \SetStartEndCondition{ }{}{}%
  \SetKwProg{Fn}{procedure}{\string:}{}
  \SetKwFunction{Range}{range}%%
  \SetKw{KwTo}{in}\SetKwFor{For}{for}{\string:}{}%
  \SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
  \SetKwFor{While}{while}{:}{fintq}%
  \AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
  \SetKwFunction{Lgc}{$lgc$}%
  \SetKwFunction{Copy}{$lgcCopy$}%
  \SetKwFunction{LCopy}{$lgcCopy$}%
  \SetKwFunction{RCopy}{$rgcCopy$}%

@@

\subsection{copying closure using liveness vs copying closures using reachability}
To compare the  effect of using reachability to  copy closures instead
of liveness we take a couple  of programs from our benchmark suite and
execute them with both. As shown in \cred{figure-add-figure-here}, the
number of  garbage collections  in both  cases are  same but  the time
taken  varies largely.  Both these  observations can  be explained  as
follows, since we approximate the  liveness of closure to the liveness
at its  creation point, the  liveness we  get is usually  grossly over
approximated.  Using this liveness may not  give too much of a benefit
compared to  reachability as  nearly everything would  be live  at its
point of  creation.  As mentioned  in section-ref, using  liveness for
closures  may  entail  multiple  traversals over  the  same  structure
consuming more time.  Therefore we believe that using reachability for
closures is  a sweet spot in  terms of performance for  liveness based
garbage collectors.
%%
%% \comment{
%% The graphs  in Figure~\ref{fig:memory-usage} show the  number of cells
%% in the heap over time for RGC and LGC---here time is measured in terms
%% of the number of cons cells allocated. In addition, they also show the
%% number of  reachable cells and the  number of cells  that are actually
%% live\footnote{This   is  statically   approximated  by   our  liveness
%%   analysis}.  Since  the programs have  different memory requirements,
%% we have tuned the size of heap for each program to ensure a reasonable
%% number of  collections. An invocation  of RGC decreases the  number of
%% cells  in heap  till  it touches  the  curve of  reachable cells.   An
%% invocation of LGC decreases the number  of heap cells to no lower than
%% the curve of live cells.

>>>>

\subsection{A liveness-based garbage collection scheme}
\label{sec:live-clo}
  \SetStartEndCondition{ }{}{}%
  \SetKwProg{Fn}{procedure}{\string:}{}
  \SetKwFunction{Range}{range}%%
  \SetKw{KwTo}{in}\SetKwFor{For}{for}{\string:}{}%
  \SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
  \SetKwFor{While}{while}{:}{fintq}%
  \AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
  \SetKwFunction{Lgc}{$lgc$}%
  \SetKwFunction{Copy}{$lgcCopy$}%
  \SetKwFunction{LCopy}{$lgcCopy$}%
  \SetKwFunction{RCopy}{$rgcCopy$}%

@@

\subsection{Copying closure using liveness vs copying closures using reachability}
To compare the  effect of using reachability to  copy closures instead
of liveness we take a couple  of programs from our benchmark suite and
execute them with both. As shown in \cred{figure-add-figure-here}, the
number of  garbage collections  in both  cases are  same but  the time
taken  varies largely.  Both these  observations can  be explained  as
follows, since we approximate the  liveness of closure to the liveness
at its  creation point, the  liveness we  get is usually  grossly over
approximated.  Using this liveness may not  give too much of a benefit
compared to  reachability as  nearly everything would  be live  at its
point of  creation.  As mentioned  in Section~\ref{sec:live-clo}, using  liveness for
closures  may  entail  multiple  traversals over  the  same  structure
consuming more time.  Therefore we believe that using reachability for
closures is  a sweet spot in  terms of performance for  liveness based
garbage collectors.
%%
%% \comment{
%% The graphs  in Figure~\ref{fig:memory-usage} show the  number of cells
%% in the heap over time for RGC and LGC---here time is measured in terms
%% of the number of cons cells allocated. In addition, they also show the
%% number of  reachable cells and the  number of cells  that are actually
%% live\footnote{This   is  statically   approximated  by   our  liveness
%%   analysis}.  Since  the programs have  different memory requirements,
%% we have tuned the size of heap for each program to ensure a reasonable
%% number of  collections. An invocation  of RGC decreases the  number of
%% cells  in heap  till  it touches  the  curve of  reachable cells.   An
%% invocation of LGC decreases the number  of heap cells to no lower than
%% the curve of live cells.
